News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-05 rupnik2015.tex
updates

In today's world, we follow news which is distributed globally. Significant events are reported by different sources and in different languages. In this work, we address the problem of tracking of events in a large multilingual stream. Within a recently developed system EventRegistry  \shortcite{Leban2014W,Leban2014I} we examine two aspects of this problem: how to compare articles in different languages and how to link collections of articles in different languages which refer to the same event.  Taking a multilingual stream and clusters of articles from each language, we compare different cross-lingual document similarity measures based on Wikipedia. This allows us to compute the similarity of any two articles regardless of language. Building on previous work, we show there are methods which scale well and can compute a meaningful similarity between articles from languages with little or no direct overlap in the training data.
Using this capability, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.
\end{abstract}

@@

The paper is organized as follows: we first provide an overview of the system as a whole (Section~\ref{sec:pipeline}), which includes a subsection that summarizes the main system requirements. We then present related work in Section~\ref{sec:related}. The related work covers work on cross-lingual document similarity as well as work on cross-lingual cluster linking. In Section~\ref{sec:crosslingual}, we introduce the problem of cross-lingual document similarity computation, followed by an overview  and description of the main cross-lingual similarity approaches we consider in detail, including a new approach based on hub languages.
%The approach in Section~\ref{sec:kmeans} serves as a baseline, and the approaches in Section~\ref{sec:CCA} and Section~\ref{sec:LSI} serve as the basis on which we proposed a new approach presented in Section~\ref{sec:hublang}
 In Section~\ref{sec:linking}, we introduce the central problem of cross-lingual linking of clusters of news articles and our approach that combines the cross-lingual similarity functions with knowledge extraction based techniques. Finally, we present and interpret the experimental results in Section~\ref{sec:evaluation} and discuss conclusions and point out several promising future directions.

@@

\noindent\textbf {Mono-lingual similarity.}
A common way of computing similarity between documents is \emph{cosine similarity},
$$sim(d_1, d_2) = \frac{\langle \phi(d_1), \phi(d_2)\rangle}{\|\phi(d_1)\| \|\phi(d_2)\|},$$
where $\langle \cdot,\cdot \rangle$ and $\|\cdot\|$ are standard inner product and Euclidean norm. When dealing with two or more languages, one could ignore the language information
and build a vector space using the union of tokens over the languages. A cosine similarity function in such a space can be useful to some extent, for example "internet" or "Obama" may appear both in a Spanish and English texts and the presence of such terms in both an English and a Spanish document would contribute to their similarity. In general however, large parts of vocabularies may not intersect, which means that given a language pair, many words in both languages cannot contribute to the similarity score, which makes the similarity function very insensitive.

@@

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we will describe several approaches to the problem of computing multilingual similarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using LSI, which
 makes the eigenvector problem computationally and numerically tractable.

@@

More formally, a \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least \textbf{two nonempty documents}. This means that in our analyses we discard strictly monolingual documents for which no cross-lingual information is available. A comparable corpus $D = {d_1, \ldots, d_s}$ is a collection of $s$ multilingual documents. By using the vector space model, we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times s}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$. See the Figure~\ref{fig:stacked_matrices} for an illustration of the introduced notation.

@@

\subsection{$k$-means}\label{sec:kmeans}

@@

\subsection{Cross-Lingual Latent Semantic Indexing}\label{sec:LSI}

@@

  Since the matrix $X$ can be large we could use an iterative method like the Lanczos~\cite{golub} algorithm with reorthogonalization to find the left singular vectors (columns of $U$) corresponding to the largest singular values. It turns out that the Lanczos method converges slowly as the gap between the leading singular values is small. Moreover, the Lanczos method is hard to parallelize. Instead, we use a randomized version of the singular value decomposition (SVD) described in \shortcite{tropp} that can be viewed as a block Lanczos method. That enables us to use parallelization and speeds up the computation considerably.

@@

\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 We now present a statistical technique to analyze data from two sources, extension of which will be presented in the next section.
%
 Canonical Correlation Analysis (CCA)~\cite{Hotelling} is a dimensionality reduction technique similar to Principal Component Analysis (PCA)~\cite{Pearson1901On}, with the additional assumption that the data consists of feature vectors that arose from two sources (two views) that share some information. Examples include: bilingual document collection ~\shortcite{mrpqr} and collection of images and captions~\shortcite{Hardoon_usingimage}. Instead of looking for linear combinations of features that maximize the variance (PCA) we look for a linear combination of feature vectors from the first view and a linear combination for the second view, that are maximally correlated.

@@

Note that the method in its original form is only applicable to two languages where an aligned set of observations is available.

@@

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English (4,212k articles)--\emph{en} (hub language), Spanish (9,686k articles)--\emph{es}, Russian (9,662k articles)--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian (136k articles)--\emph{sl}, Piedmontese (59k articles)--\emph{pms}, Waray-Waray (112k articles)--\emph{war} (all with about 2 million native speakers), Creole (54k articles)--\emph{ht} (8 million native speakers), and Hindi (97k articles)--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs \footnote{Such documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.}) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors. The IDF scores are computed for each language with the exception of English based on its aligned documents with the English Wikipedia. The English language IDF scores are based on all English documents for which aligned Spanish documents exist.

@@

 The hub CCA outperforms other methods in this sense in terms of precision, which is consistent with its AMRR performance, see Figure~\ref{pic:AMRR}.
It is worth noting that added accuracy is also beneficial when we wish to choose good representatives of clusters. Therefore, we decided to use hub CCA in the cluster linking component.

@@

\subsection{Remarks on the repeatability of Experiments}
We have made  both the code and data that were used in the experiments publicly available at \url{https://github.com/rupnikj/jair_paper.git}.
The manually labelled dataset used in the evaluation of event linking is available at in the ``dataset`` subfolder of the github repository. The included archive contains two folders: ``positive'' and ``negative'', where the first folder includes examples of cluster pairs in two languages that represent the same event and the second folder contains pairs of clusters in two languages that do not represent different events. Each example is a JSON file that contains at the top level information about a pair of clusters (including text of the articles) as well as a set of ``meta'' attributes, that correspond to features described in Section \ref{algo:features}.

>>>>

In today's world, we follow news which is distributed globally. Significant events are reported by different sources and in different languages. In this work, we address the problem of tracking of events in a large multilingual stream. Within a recently developed system Event Registry  \shortcite{Leban2014W,Leban2014I} we examine two aspects of this problem: how to compare articles in different languages and how to link collections of articles in different languages which refer to the same event.  Taking a multilingual stream and clusters of articles from each language, we compare different cross-lingual document similarity measures based on Wikipedia. This allows us to compute the similarity of any two articles regardless of language. Building on previous work, we show there are methods which scale well and can compute a meaningful similarity between articles from languages with little or no direct overlap in the training data.
Using this capability, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.
\end{abstract}

@@

The paper is organized as follows: we first provide an overview of the system as a whole in Section~\ref{sec:pipeline}, which includes a subsection that summarizes the main system requirements. We then present related work in Section~\ref{sec:related}. The related work covers work on cross-lingual document similarity as well as work on cross-lingual cluster linking. In Section~\ref{sec:crosslingual}, we introduce the problem of cross-lingual document similarity computation, which is followed by an overview  and description of the main cross-lingual similarity approaches we consider in detail, including a new approach based on hub languages.
%The approach in Section~\ref{sec:kmeans} serves as a baseline, and the approaches in Section~\ref{sec:CCA} and Section~\ref{sec:LSI} serve as the basis on which we proposed a new approach presented in Section~\ref{sec:hublang}
 In Section~\ref{sec:linking}, we introduce the central problem of cross-lingual linking of clusters of news articles and our approach that combines the cross-lingual similarity functions with knowledge extraction based techniques. Finally, we present and interpret the experimental results in Section~\ref{sec:evaluation} and discuss conclusions and point out several promising future directions.

@@

\noindent\textbf {Mono-lingual similarity.}
A common way of computing similarity between documents is \emph{cosine similarity},
$$sim(d_1, d_2) = \frac{\langle \phi(d_1), \phi(d_2)\rangle}{\|\phi(d_1)\| \|\phi(d_2)\|},$$
where $\langle \cdot,\cdot \rangle$ and $\|\cdot\|$ are standard inner product and Euclidean norm. When dealing with two or more languages, one could ignore the language information
and build a vector space using the union of tokens over the languages. A cosine similarity function in such a space can be useful to some extent, for example "internet" or "Obama" may appear both in Spanish and English texts and the presence of such terms in both an English and a Spanish document would contribute to their similarity. In general however, large parts of vocabularies may not intersect, which means that given a language pair, many words in both languages cannot contribute to the similarity score, which makes the similarity function very insensitive to the data.

@@

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we will describe several approaches to the problem of computing multilingual similarities introduced in Section~\ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in Section~\ref{sec:kmeans}, a standard approach based on singular value decomposition in Section~\ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in Section~\ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in Section~\ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English for Wikipedia), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using CL-LSI, which
 makes the eigenvector problem computationally and numerically tractable.

@@

More formally, a \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least \textbf{two nonempty documents}. This means that in our analysis we discard strictly monolingual documents for which no cross-lingual information is available. A comparable corpus $D = {d_1, \ldots, d_s}$ is a collection of $s$ multilingual documents. By using the vector space model, we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times s}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$. See the Figure~\ref{fig:stacked_matrices} for an illustration of the introduced notation.

@@

We will now describe four models to cross-lingual similarity computation in the next sections.
\subsection{$k$-means}\label{sec:kmeans}

@@

In the next section we will describe our second approach.
\subsection{Cross-Lingual Latent Semantic Indexing}\label{sec:LSI}

@@

  Since the matrix $X$ can be large we could use an iterative method like the Lanczos algorithm with reorthogonalization~\cite{golub} to find the left singular vectors (columns of $U$) corresponding to the largest singular values. It turns out that the Lanczos method converges slowly as the gap between the leading singular values is small. Moreover, the Lanczos method is hard to parallelize. Instead, we use a randomized version of the SVD \shortcite{tropp} that can be viewed as a block Lanczos method. That enables us to use parallelization and speeds up the computation considerably.

@@

In the next section we will introduce our third approach.
\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 We now present a statistical technique to analyze data from two sources, extension of which will be presented in the next section.
%
 Canonical Correlation Analysis (CCA)~\cite{Hotelling} is a dimensionality reduction technique similar to Principal Component Analysis (PCA)~\cite{Pearson1901On}, with the additional assumption that the data consists of feature vectors that arose from two sources (two views) that share some information. Examples include: bilingual document collection ~\shortcite{mrpqr} and collection of images and captions~\shortcite{Hardoon_usingimage}. Instead of looking for linear combinations of features that maximize the variance (PCA) we look for a linear combination of feature vectors from the first view and a linear combination for the second view, that are maximally correlated.

@@

Note that the method in its original form is only applicable to two languages where an aligned set of observations is available. The next section will describe a scalable an extension of CCA to more than two languages.

@@

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English (4,212k articles)--\emph{en} (hub language), Spanish (9,686k articles)--\emph{es}, Russian (9,662k articles)--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian (136k articles)--\emph{sl}, Piedmontese (59k articles)--\emph{pms}, Waray-Waray (112k articles)--\emph{war} (all with about 2 million native speakers), Creole (54k articles)--\emph{ht} (8 million native speakers), and Hindi (97k articles)--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (referred to as stubs\footnote{Such documents are typically of low value as a linguistic resource. Examples include the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.}) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors. The IDF scores are computed for each language with the exception of English based on its aligned documents with the English Wikipedia. The English language IDF scores are based on all English documents for which aligned Spanish documents exist.

@@

The hub CCA attains higher precision and classification accuracy on the task of linking small cluster pairs than the other methods, while LSI is slightly better on linking large cluster pairs. The gain in precision of LSI over hub CCA on linking large clusters is much smaller than the gain in precision of hub CCA over LSI on linking small clusters. For that reason we decided to use hub CCA as the similarity computation component in our system.

@@

\subsection{Remarks on the repeatability of experiments}
We have made  both the code and data that were used in the experiments publicly available at \url{https://github.com/rupnikj/jair_paper.git}.
The manually labelled dataset used in the evaluation of event linking is available at in the ``dataset`` subfolder of the github repository. The included archive contains two folders: ``positive'' and ``negative'', where the first folder includes examples of cluster pairs in two languages that represent the same event and the second folder contains pairs of clusters in two languages that do not represent different events. Each example is a JSON file that contains at the top level information about a pair of clusters (including text of the articles) as well as a set of ``meta'' attributes, that correspond to features described in Section \ref{algo:features}.
